grouping and sumarizing

Sumarizing the median life expectancy

```{r}
gapminder %>%
  summarize(medianLifeExp = median(lifeExp))

#Filter for 1952 then summarize the median life expectancy

gapminder %>%
  filter(year == 1957) %>%
  summarize(medianLifeExp = median(lifeExp))

gapminder %>%
  filter(year == 1957) %>%
  summarize(medianLifeExp = median(lifeExp), maxGdpPerCap = max(gdpPercap))


```

Summarizin by year
```{r}
gapminder %>%
  group_by(year) %>%
  summarize(medianLifeExp = median(lifeExp), maxGdpPerCap = max(gdpPercap))


```
summary by continent

```{r}

gapminder %>%
  filter(year == 1957) %>%
  group_by(continent) %>%
  summarize(medianLifeExp = median(lifeExp), maxGdpPerCap = max(gdpPercap)) %>%
  arrange(desc(maxGdpPerCap))


```

summarizin by continent and year

```{r}
gapminder %>%
  group_by(continent, year) %>%
  summarize(medianLifeExp = median(lifeExp), maxGdpPerCap = max(gdpPercap))
  
  
```

visualizing median life expectancy over time

```{r}

by_year <- gapminder %>%
  group_by(year) %>%
  summarize(medianLifeExp = median(gdpPercap))

by_year

ggplot(by_year, aes(x=year,y=medianLifeExp)) +
  geom_line()


```
```{r}
by_year_continent <- gapminder %>%
  group_by(year, continent) %>%
  summarize(medianGdpPerCap = median(gdpPercap))

by_year_continent

ggplot(by_year_continent, aes(x=year,y=medianGdpPerCap, color=continent))+
  geom_line()

```

Bar plots

visualizing median GDP per capita by continent

```{r}

by_continent <- gapminder %>%
  filter(year == 1952) %>%
  group_by(continent) %>%
  summarize(medianGdpPerCap = median(gdpPercap))

by_continent

ggplot(by_continent, aes(x=continent,y=medianGdpPerCap))+
  geom_col()

```
visualizing GDP per capita by country in oceania

```{r}

oceania_1952 <- gapminder%>%
  filter(continent == "Oceania", year == 1952)

oceania_1952

ggplot(oceania_1952, aes(x=country, y=gdpPercap))+
  geom_col()
```
Histogram

visualizing population

```{r}
library("gapminder")
library("ggplot2")
library("dplyr")

gapminder

gapminder_1952 <- gapminder %>%
  filter(year==1952) %>%
  mutate(pop_by_mil = pop/1000000)


ggplot(gapminder_1952, aes(x=pop_by_mil)) +
  geom_histogram(bins = 50)


```

visualiziing population con x-aaxis on a log scale
```{r}
gapminder_1952 <- gapminder %>%
  filter(year==1952)

ggplot(gapminder_1952, aes(x=pop)) +
  geom_histogram(bins = 50) +
  scale_x_log10()


```

Boxplots

comparing GDP per capita across continents

```{r}
gapminder_1952 <- gapminder %>%
  filter(year==1952)

ggplot(gapminder, aes(x=continent, y = gdpPercap))+
  geom_boxplot()+
  scale_y_log10()

```
adding title to your graph

```{r}
ggplot(gapminder, aes(x=continent, y = gdpPercap))+
  geom_boxplot()+
  scale_y_log10()+
  ggtitle("Comparing GDP per capita across continents")


```

Data manipulation with  dplyr

```{r}
library(tidyverse)
counties <- read.csv('counties.csv')
glimpse(counties)

counties


```
Selecting columns

```{r}

counties %>%
  select(state, county, population, men, women)

counties_selected <- counties %>%
  select(state, county, population, private_work, public_work, self_employed)

counties_selected %>%
  arrange(desc(public_work))

counties_selected <- counties %>%
  select(state, county, population)

counties_selected %>%
  filter(population > 1000000)

counties_selected <- counties %>%
  select(state, county, population, private_work, public_work, self_employed)

counties_selected%>%
  filter(state=="Texas", population > 10000)%>%
  arrange(desc(private_work))

```
select, mutate, filter and arrange


```{r}

counties %>%
  select(state, county, population, men, women) %>%
  mutate(proportion_men = men/population) %>%
  filter(population > 10000) %>%
  arrange(desc(proportion_men))

```
the verb count
```{r}

select_counties <- counties %>%
  select(state, region, county, population, citizens)

counties_selected <- counties %>%
  count(region, sort = TRUE) %>%
  arrange(desc(n))

```
counting citizens by state

```{r}
counties_selected <- counties %>%
  select(county, region, state, population, citizens)

#agrupa con state cuenta citizens
counties_selected %>%
  count(state, wt=citizens)%>%
  arrange(desc(n))

```
Mutating and counting

```{r}
counties_selected <- counties %>%
  select(county, region, state, population,walk)

counties_selected %>%
  mutate(population_walk = population*walk/100)%>%
  count(state,wt=population_walk,sort = TRUE)


```
summarizing

```{r}

counties_selected <- counties %>%
  select(county, population, income, unemployment)

counties_selected %>%
  summarize(min_population = min(population), max_unemployment = max(unemployment), averagen_income = mean(income))


```
summarizing by state

```{r}

counties_selected <- counties %>%
  select(state,county, population, land_area)

counties_selected %>%
  group_by(state) %>%
  summarize(total_area = sum(land_area), total_population = sum(population))%>%
  mutate(density= total_population/total_area)%>%
  arrange(desc(density))

counties_selected <- counties %>%
  select(region, state, county, population)

counties_selected %>%
  group_by(region, state)%>%
  summarize(total_pop = sum(population))%>%
  summarize(average_pop=mean(total_pop), median_pop = median(total_pop))

```

Top_n verb

```{r}
counties_selected <- counties %>%
  select(region, state, county, metro, population, walk)

counties_selected %>%
  group_by(region)%>%
  top_n(1,walk)

counties_selected <- counties %>%
  select(region, state, county, population, income)

counties_selected %>%
  group_by(region, state)%>%
  summarize(average_income = mean(income))%>%
  top_n(1,average_income)

```



```{r}
counties_selected <- counties %>%
  select(state, metro, population)

counties_selected %>%
  group_by(state, metro) %>%
  summarize(total_pop = sum(population))%>%
  top_n(1,total_pop) %>%
  ungroup() %>%
  count(metro)


```

selecting columns

```{r}

counties %>%
  select(state, county, population, professional:production) %>%
  arrange(desc(service))

counties %>%
  select(state, county, population, ends_with("work"))%>%
  filter(public_work >= 50)

counties %>%
  select(state, county, population, starts_with("income"))

counties %>%
  select(state, county, population, contains("at"))

```

renaming column after count

```{r}
library(dplyr)
counties%>%
  count(state) %>%
  rename(num_counties = n)

counties %>%
  select(state,county,poverty_rate = poverty)

```
Using transmute

```{r}
counties %>%
  transmute(state, county, population,density=population/land_area)%>%
  filter(population >= 1000000) %>%
  arrange(desc(density))


```
choosing among the four verbs

```{r}
##change the name of uneployment column

counties %>%
  rename(unemployment_rate=unemployment)

#keep the state and punty coluns, and the columns containing poverty

counties%>%
  select(state, county, contains("poverty"))

#calculate the function_Women column without dropping the other columns

counties%>%
  mutate(fraction_women = women/population)

#keep only the state, county, and employement_rate columns

counties %>%
  transmute(state, county, employment_rate = employed/population)


```




```{r}



```

