```{r}

library("gapminder")
library("dplyr")

gapminder



```
filteriong for one year

```{r}

gapminder %>%
  filter(year==1957, country=="China")



```

arranging observations by life expectancy
```{r}

gapminder %>%
  arrange(lifeExp)

#Sort in descending order of lifeExp

gapminder %>%
  arrange(desc(lifeExp))

```

filtering and arranging
```{r}

gapminder %>%
  filter(year == 1957) %>%
  arrange(desc(pop))

```
Using mutate to change or create a column

```{r}
#use mutate to change lifeExp to be in months
gapminder %>%
  mutate(lifeExp = 12*lifeExp)

#use mutate to create a new column called lifeExpMonths
gapminder %>%
  mutate(lifeExpMonths = 12 * lifeExp)

```
Combining filter, mutate and arrange

```{r}

gapminder %>%
  filter(year == 2007) %>%
  mutate(lifeExpMonths = 12 * lifeExp) %>%
  arrange(desc(lifeExpMonths))

```
visualizing with ggplot2
```{r}
library(ggplot2)

gapminder_1952 <- gapminder %>%
  filter(year==1952)

ggplot(gapminder_1952, aes(x=pop,y=gdpPercap)) +
  geom_point()

```

```{r}

ggplot(gapminder_1952, aes(x=pop,y=lifeExp)) +
  geom_point()

```

Potting the x-axis on a log scale
```{r}

gapminder_1952 <- gapminder %>%
  filter(year==1952)

ggplot(gapminder_1952, aes(x=pop,y=lifeExp)) +
  geom_point() +
  scale_x_log10()

```

putting the x-and and y-axes on a log scale
```{r}

ggplot(gapminder_1952, aes(x=pop, y=gdpPercap))+
  geom_point()+
  scale_x_log10()+
  scale_y_log10()

```

adding color to a scatter plot
```{r}

ggplot(gapminder_1952, aes(x=pop, y=lifeExp, color=continent))+
  geom_point() +
  scale_x_log10()

```

adding size and color to a plot

```{r}
ggplot(gapminder_1952, aes(x=pop, y=lifeExp, color=continent, size=gdpPercap))+
  geom_point() +
  scale_x_log10()

```
Faceting

creating a subgraph for each continent

```{r}
gapminder_1952
ggplot(gapminder_1952, aes(x=pop, y=lifeExp))+
  geom_point() +
  scale_x_log10() +
  facet_wrap(~ continent)

```

faceting by year

```{r}

gapminder_1952
ggplot(gapminder, aes(x=gdpPercap, y=lifeExp, color=continent, size=pop))+
  geom_point() +
  scale_x_log10() +
  facet_wrap(~ year)

```


